Unifying Lower Bounds on Prediction Dimension of Consistent Convex Surrogates

### 2021-05-18 cvx-flats-neurips21/cvx-flats.tex
Turning comments back on and adding opinions about things going in Appendix

%\usetikzlibrary{calc}
\newcommand{\Comments}{0}
\newcommand{\mynote}[2]{\ifnum\Comments=1\textcolor{#1}{#2}\fi}
\newcommand{\mytodo}[2]{\ifnum\Comments=1%
	\todo[linecolor=#1!80!black,backgroundcolor=#1,bordercolor=#1!80!black]{#2}\fi}
\newcommand{\raf}[1]{\mynote{green}{[RF: #1]}}
\newcommand{\raft}[1]{\mytodo{green!20!white}{RF: #1}}
\newcommand{\jessie}[1]{\mynote{purple}{[JF: #1]}}
\newcommand{\jessiet}[1]{\mytodo{purple!20!white}{JF: #1}}
\newcommand{\bo}[1]{\mynote{blue}{[Bo: #1]}}
\newcommand{\botodo}[1]{\mytodo{blue!20!white}{[Bo: #1]}}
\newcommand{\btw}[1]{}%\mytodo{orange!80!white}{BTW: #1}}
\newcommand{\discuss}[1]{\mynote{cyan!20!white}{#1}}
\ifnum\Comments=1               % fix margins for todonotes
\setlength{\marginparwidth}{1in}
\fi

@@

Instead of a target loss $\ell$, one may want to learn a target property, i.e. a conditional statistic such as the expected value, variance, or entropy.
In this case, following the tradition in the statistics literature on conditional estimation~\citep{gyorfi2006distribution,fan1998efficient,ruppert1997local},
we formalize consistency as converging to the correct conditional estimates of the property.
Convergence is measured by functions $\propdis(r, p)$ that formalize how close $r$ is to ``correct'' for conditional distribution $p$.
In particular we should have $\propdis(r,p) = 0 \iff r \in \gamma(p)$.
\btw{Bo: Would be nice to give some natural special cases: for a finite property, $\propdis(r,p) = \Ind{r \in \gamma(p)}$, and for single-valued properties with a distance metric on $\R$, $\propdis(r,p) = \text{dist}(r, \gamma(p))$.}

@@

%Lemma~\ref{lem:consistent-loss-implies-prop} confirms that consistency with respect to a target property is a strictly broader notion than with respect to a target loss.
Because each target loss in $\L$ elicits some property, but not all target properties can be elicited by a loss (e.g. the variance), consistency with respect to a property is the strictly broader notion.
This points to indirect elicitation as a natural necessary condition for consistency, as formalized in Theorem \ref{thm:consistent-implies-indir-elic}.

@@

\section{Prediction Dimension of Consistent Convex Surrogates}\label{sec:char-convex}
We now turn to the question of bounding the prediction dimension of a consistent convex surrogate.
From Theorem \ref{thm:consistent-implies-indir-elic}, given a target property $\gamma$ or loss $\ell$ with $\gamma = \prop{\ell}$, this task reduces to lower bounding the prediction dimension of a convex surrogate indirectly eliciting $\gamma$.
We now explore two tools, Corollaries~\ref{cor:Pcodim-flat-single-val-prop} and~\ref{cor:Pcodim-flat-elic-relint-prop}, for proving such convex elicitation lower bounds.
The key idea, crystallized from the proofs of \citet[Theorem 16]{ramaswamy2016convex} and \citet[Theorem~9]{agarwal2015consistent}, is to consider a particular distribution~$p$ and surrogate prediction $u \in \reals^d$ with is optimal for $p$.
Theorem~\ref{lem:convex-flats-inf-dim} will show that if $d$ is small, then the level set $\{p \in \P : u \in \argmin_{u'} \exploss{L}{u'}{p}\}$ must be large; in fact, it must roughly contain a high-dimensional \emph{flat}.
By definition of indirect elicitation, there is some level set $\gamma_r$ (where $u$ is linked to $r$) containing this flat as well.
The use of this result is to leverage the contrapositive: if $\gamma$ has a level set intricate enough to not contain any high-dimensional flats, then $\gamma$ cannot have a low-dimensional consistent surrogate.

@@

\begin{lemma}
  \label{lem:convex-flats-inf-dim}
  Let $\Gamma:\P \toto \reals^d$ be (directly) elicited by $L \in \Lcvx_d$ for some $d\in\mathbb{N}$.
  Let $\Y$ be either a finite set, or $\Y = \reals$, in which case we assume each $p\in\P$ admits a Lebesgue density supported on the same set for all $p\in\P$.%
  \footnote{This assumption is largely for technical convenience, to ensure that $\V_{u,p}$ does not depend on $p$.
    Any such assumption would suffice, and we suspect even that condition can be relaxed.}
  For all $u\in\range\Gamma$ and $p\in\Gamma_u$, there is some $V_{u,p}:\Y\to\reals^d$ such that $p \in \zeros{V_{u,p}} \subseteq \Gamma_u$.
\end{lemma}
\begin{proof}
  As $L$ is convex and elicits $\Gamma$, we have $u \in \Gamma(p) \iff \vec 0 \in \partial \exploss{L}{u}{p}$. 
  We proceed in two cases, depending on $|\Y|$.

@@

\section{Illustration: geometric intuition for the four quadrants}
We can apply Lemma~\ref{lem:convex-flats-inf-dim} to construct lower bounds on convex consistency dimension for targets across all four quadrants by Corollaries~\ref{cor:Pcodim-flat-elic-relint-prop} (for discrete prediction problems; Q1, Q2) and~\ref{cor:Pcodim-flat-single-val-prop} (for continuous estimation problems; Q3, Q4).
This is demonstrated in Table~\ref{tab:illustrate-quadrant-apps}.

@@

\section{Breadth of framework: Discrete-valued predictions}\label{sec:discrete-predictions-refactored}
In discrete prediction settings (Quadrants 1 and 2), we can apply Lemma~\ref{lem:convex-flats-inf-dim} to observe lower bounds on convex prediction dimension that subsume previous bounds given by~\citet{ramaswamy2016convex}.
Their lower bound 
%(the only non-trivial one to our knowledge) 
is given by the co-dimension of the \emph{feasible subspace dimension} $\Sc_\C(p)$ of a convex set $\C$ at a given distribution $p$ such that $p \in \C$, which is loosely the highest dimension subspace of $\C$ containing a neighborhood around $p$.
The feasible subspace dimension uses ``local'' information of the property to construct lower bounds, while $d$-flats allow us to also use ``global'' information, while subsuming the previous results about local information.

@@

In Appendix~\ref{sec:finite-calib}, we show examples where the $d$-flats bound both matches the feasible subspace dimension bound and strictly improves it in order to emphasize that our result is strictly broader.

@@

\paragraph{Example: Variance.}
As a warm-up, let us see how to show $\eliccvx(\Var)=2$, meaning the lowest dimension of a convex loss to estimate conditional variance is 2. %~\citep{osband1985providing,lambert2018elicitation}.
This lower bound will follow from Theorem~\ref{thm:bayes-risk-lower-bound} using that variance is the Bayes risk of squared loss $L(r,y) = (r-y)^2$, which elicits the mean $\Gamma(p) = \E_p Y$.
Interestingly, while perhaps intuitively obvious, even this simple result is novel.
In particular, the well-known fact that the variance is not elicitable does not yield a lower bound of 2, as it does not rule out the variance being a link of a real-valued convex-elicitable property; cf.~\citet[Remark 1]{frongillo2020elicitation}.

@@

\section{Conclusions and future work}\label{sec:conclusions}
In this work, we show that indirect property elicitation can be a powerful necessary condition for the existence of a consistent surrogate loss (Theorem~\ref{thm:consistent-implies-indir-elic}).
Furthermore, we introduce a new lower bound (Corollaries~\ref{cor:Pcodim-flat-single-val-prop} and~\ref{cor:Pcodim-flat-elic-relint-prop}) on convex consistency dimension that is generally applicable and extends previous results from both the discrete (Corollary~\ref{cor:fsd-bound}) and continuous (Corollaries~\ref{cor:variance} and \ref{cor:spectral-risks}) estimation settings.

>>>>

%\usetikzlibrary{calc}
\newcommand{\Comments}{1}
\newcommand{\mynote}[2]{\ifnum\Comments=1\textcolor{#1}{#2}\fi}
\newcommand{\mytodo}[2]{\ifnum\Comments=1%
	\todo[linecolor=#1!80!black,backgroundcolor=#1,bordercolor=#1!80!black]{#2}\fi}
\newcommand{\raf}[1]{\mynote{green}{[RF: #1]}}
\newcommand{\raft}[1]{\mytodo{green!20!white}{RF: #1}}
\newcommand{\jessie}[1]{\mynote{purple}{[JF: #1]}}
\newcommand{\jessiet}[1]{\mytodo{purple!20!white}{JF: #1}}
\newcommand{\bo}[1]{\mynote{blue}{[Bo: #1]}}
\newcommand{\botodo}[1]{\mytodo{blue!20!white}{[Bo: #1]}}
\newcommand{\btw}[1]{}%\mytodo{orange!80!white}{BTW: #1}}
\newcommand{\discuss}[1]{\mynote{cyan!20!white}{#1}}
\ifnum\Comments=1               % fix margins for todonotes
\setlength{\marginparwidth}{1in}
\fi

@@

Instead of a target loss $\ell$, one may want to learn a target property, i.e. a conditional statistic such as the expected value, variance, or entropy.
In this case, following the tradition in the statistics literature on conditional estimation~\citep{gyorfi2006distribution,fan1998efficient,ruppert1997local},
we formalize consistency as converging to the correct conditional estimates of the property.
Convergence is measured by functions $\propdis(r, p)$ that formalize how close $r$ is to ``correct'' for conditional distribution $p$.
In particular we should have $\propdis(r,p) = 0 \iff r \in \gamma(p)$.
\btw{Bo: Would be nice to give some natural special cases: for a finite property, $\propdis(r,p) = \Ind{r \in \gamma(p)}$, and for single-valued properties with a distance metric on $\R$, $\propdis(r,p) = \text{dist}(r, \gamma(p))$.}
\jessie{Move this statement to appendix, and modify above paragraph to give hand-waving description.}

@@

\jessie{Comment on how this lets us ``shift'' Q1 to Q2, and Q3 to Q4.  This is a subtlety missing in Table~\ref{tab:quadrants-simplices} too.}
%Lemma~\ref{lem:consistent-loss-implies-prop} confirms that consistency with respect to a target property is a strictly broader notion than with respect to a target loss.
Because each target loss in $\L$ elicits some property, but not all target properties can be elicited by a loss (e.g. the variance), consistency with respect to a property is the strictly broader notion.
This points to indirect elicitation as a natural necessary condition for consistency, as formalized in Theorem \ref{thm:consistent-implies-indir-elic}.

@@

\section{Prediction Dimension of Consistent Convex Surrogates}\label{sec:char-convex}
We now turn to the question of bounding the prediction dimension of a consistent convex surrogate.
From Theorem \ref{thm:consistent-implies-indir-elic}, given a target property $\gamma$ or loss $\ell$ with $\gamma = \prop{\ell}$, this task reduces to lower bounding the prediction dimension of a convex surrogate indirectly eliciting $\gamma$.
We now explore two tools, Corollaries~\ref{cor:Pcodim-flat-single-val-prop} and~\ref{cor:Pcodim-flat-elic-relint-prop}, for proving such convex elicitation lower bounds.
The key idea, crystallized from the proofs of \citet[Theorem 16]{ramaswamy2016convex} and \citet[Theorem~9]{agarwal2015consistent}, is to consider a particular distribution~$p$ and surrogate prediction $u \in \reals^d$ with is optimal for $p$.
Theorem~\ref{lem:convex-flats-inf-dim} will show that if $d$ is small, then the level set $\{p \in \P : u \in \argmin_{u'} \exploss{L}{u'}{p}\}$ must be large; in fact, it must roughly contain a high-dimensional \emph{flat}.
By definition of indirect elicitation, there is some level set $\gamma_r$ (where $u$ is linked to $r$) containing this flat as well.
The use of this result is to leverage the contrapositive: if $\gamma$ has a level set intricate enough to not contain any high-dimensional flats, then $\gamma$ cannot have a low-dimensional consistent surrogate.
\jessie{Edit this paragraph if Cors 2 and 3 move to the Appendix}

@@

\begin{lemma}
  \label{lem:convex-flats-inf-dim}
  Let $\Gamma:\P \toto \reals^d$ be (directly) elicited by $L \in \Lcvx_d$ for some $d\in\mathbb{N}$.
  Let $\Y$ be either a finite set, or $\Y = \reals$, in which case we assume each $p\in\P$ admits a Lebesgue density supported on the same set for all $p\in\P$.%
  \footnote{This assumption is largely for technical convenience, to ensure that $\V_{u,p}$ does not depend on $p$.
    Any such assumption would suffice, and we suspect even that condition can be relaxed.}
  For all $u\in\range\Gamma$ and $p\in\Gamma_u$, there is some $V_{u,p}:\Y\to\reals^d$ such that $p \in \zeros{V_{u,p}} \subseteq \Gamma_u$.
  \jessie{Change back to Theorem}
\end{lemma}
\begin{proof}
  As $L$ is convex and elicits $\Gamma$, we have $u \in \Gamma(p) \iff \vec 0 \in \partial \exploss{L}{u}{p}$. 
  We proceed in two cases, depending on $|\Y|$.

@@

\section{Illustration: geometric intuition for the four quadrants}
\jessiet{At this rate, doesn't need to be in a section header of its own?}
We can apply Lemma~\ref{lem:convex-flats-inf-dim} to construct lower bounds on convex consistency dimension for targets across all four quadrants by Corollaries~\ref{cor:Pcodim-flat-elic-relint-prop} (for discrete prediction problems; Q1, Q2) and~\ref{cor:Pcodim-flat-single-val-prop} (for continuous estimation problems; Q3, Q4).
This is demonstrated in Table~\ref{tab:quadrants-simplices}.
These illustrations are simply intended to show how Lemma~\ref{lem:convex-flats-inf-dim} can be applied to results in all four quadrants of problem types-- making it the first lower bound to our knowledge to do so.

@@

\section{Breadth of framework: Discrete-valued predictions}\label{sec:discrete-predictions-refactored}
In discrete prediction settings (Quadrants 1 and 2), we can apply Lemma~\ref{lem:convex-flats-inf-dim} to observe lower bounds on convex prediction dimension that subsume previous bounds given by~\citet{ramaswamy2016convex}.
Their lower bound 
%(the only non-trivial one to our knowledge) 
is given by the co-dimension of the \emph{feasible subspace dimension} $\Sc_\C(p)$ of a convex set $\C$ at a given distribution $p$ such that $p \in \C$, which is loosely the highest dimension subspace of $\C$ containing a neighborhood around $p$.
The feasible subspace dimension uses ``local'' information of the property to construct lower bounds, while $d$-flats in Lemma~\ref{lem:convex-flats-inf-dim} allow us to also use ``global'' information, while subsuming the previous results about local information.

@@

In Appendix~\ref{sec:finite-calib}, we show examples where the $d$-flats bound both matches the feasible subspace dimension bound and strictly improves it in order to emphasize that our result is strictly broader than previous bounds for Quadrant 1 from~\citet{ramaswamy2016convex}.

@@

\jessie{Suggest cutting this for space, since it's in the table.}
\paragraph{Example: Variance.}
As a warm-up, let us see how to show $\eliccvx(\Var)=2$, meaning the lowest dimension of a convex loss to estimate conditional variance is 2. %~\citep{osband1985providing,lambert2018elicitation}.
This lower bound will follow from Theorem~\ref{thm:bayes-risk-lower-bound} using that variance is the Bayes risk of squared loss $L(r,y) = (r-y)^2$, which elicits the mean $\Gamma(p) = \E_p Y$.
Interestingly, while perhaps intuitively obvious, even this simple result is novel.
In particular, the well-known fact that the variance is not elicitable does not yield a lower bound of 2, as it does not rule out the variance being a link of a real-valued convex-elicitable property; cf.~\citet[Remark 1]{frongillo2020elicitation}.

@@

\section{Conclusions and future work}\label{sec:conclusions}
In this work, we show that indirect property elicitation can be a powerful necessary condition for the existence of a consistent surrogate loss (Theorem~\ref{thm:consistent-implies-indir-elic}).
Furthermore, we introduce a new lower bound (Corollaries~\ref{cor:Pcodim-flat-single-val-prop} and~\ref{cor:Pcodim-flat-elic-relint-prop}
\jessie{Lemma~\ref{lem:convex-flats-inf-dim}}) on convex consistency dimension that is generally applicable and extends previous results from both the discrete (Corollary~\ref{cor:fsd-bound}) and continuous (Corollaries~\ref{cor:variance} and \ref{cor:spectral-risks}) estimation settings.
